---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# kmodesFlow

<!-- badges: start -->
<!-- badges: end -->

The *kmodesFlow* package provides convenience functions to accompany the *klaR::kmodes()* function, improving overall workflow with kmodes modeling. It allows users to easily fit multiple models, assess model fit, and examine results with automated visualizations.

## Installation

You can install the development version of kmodesFlow from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Sdbock/kmodesFlow")
```

## Tutorial

The workhorse of the *kmodesFlow* package is the *fit_models()* function. To fit 8 models, with 1 through 8-k solutions, simply specify "1:8" as the number for k: 

```{r eval = FALSE}
library(kmodesFlow)

fit_models(
  data = data,
  k = 1:8,
  seed = 1234
)
```

The *fit_models()* returns a tibble with the number of rows equal to the number of models specified. The columns contain information about cluster assignment for respondents, model fit, and visualization of cluster profiles. To examine the modal category for each variable within clusters, print the *table_cluster_modes* column for the relevant model. I suggest using the *pluck()* function to identify and print the desired row and column. To look at the modes across clusters for the $k = 3$ model: 

```{r eval = FALSE}
pluck(model_output, "table_cluster_modes", 3)
```

There is also a heat map of the distributions of variable levels within each cluster, which can be accessed in the "table_attribute_distribution" column:

```{r eval = FALSE}
pluck(model_output, "table_attribute_distribution", 3)
```

Finally, the  *plot_elbow()* function takes a *fit_models()* output and returns a plot of the sum of within-cluster differences across each level of $k$. This "elbow" plot is a common tool in determining the optimal number of clusters.

```{r eval = FALSE}
plot_elbow(model_output)
```


See Demo for a fuller overview of the *kmodesFlow* package.


